---
title: "Class work practice"
author: "Jonathan Agyeman"
date: "26 September 2015"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Performing analysis on gapminder

I first begin by calling the gapminder data and then exploring the data.

```{r, echo = TRUE}
library(gapminder)
str(gapminder) # look at structure of data
head(gapminder) #call top six rows
tail(gapminder) #call last six rows
gapminder[sample(nrow(gapminder),10),] # sample 10 rows randomly
names(gapminder) # variable names of data set
dim(gapminder) # dimension of data set
summary(gapminder) # a quick summary of all the variables in the data set
class(gapminder) # class of the data set
class(gapminder$continent) # the class of continent variable
nlevels(gapminder$continent) # number of levels the continent variable has
levels(gapminder$continent) # names of levels of continent variable
table(gapminder$continent) # count of the continent variable under various levels
```

*************************************************************************************
## Exploring data using plots

Now I explore the variables in the data set using r plots functions.

```{r ,echo= TRUE}
plot(lifeExp ~ year, gapminder) # This plots lifeExp against year without attaching data set.

plot(gapminder$year, gapminder$lifeExp) # this gives the same plot

plot(lifeExp ~ gdpPercap, gapminder) 

plot(lifeExp ~ log(gdpPercap), gapminder) # transforming the gdpPercap data for a better plot

hist(gapminder$lifeExp) # histogram of the lifeExp variable for individual summaries

barplot(table(gapminder$continent)) # barplot of the continents
```
----------

## Exploring data using ggplot2

I now try out the plot functions of the ggplot2

```{r}
library(ggplot2) # call ggplot2 package

p <- ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) # call the function for variable are being used in the plot.

p + geom_point() # plots a scatter plot.

p + scale_x_log10() + geom_point() # plots the scatter plot and logs the x varaible

p <- p + scale_x_log10() 

p + geom_point(aes(color = continent)) # gives a plot with color coding for the continents

p + geom_point(alpha = 0.05)# alpha is used to control the transparency of the points in the plot. The lower the value of alpha the more transparent the points.

p + geom_point(aes(color=continent), alpha = (1), size = 2) # we can determine the color, the transparency and the size of the points in the plot.

```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
